---
title: "Latent growth models"
format: html
editor: visual
---

# Read in data

```{r}
library(here)
library(lavaan)
library(tidyverse)
library(broom)
data.wide <- read_csv(here('Generated Datasets', 'data_wide.rds'))
```

# Generic model functions

## Linear

```{r}

linear.model.generic <- '

myIntercept =~ 1*Var.1 +
              1*Var.2 +
              1*Var.3 +
              1*Var.4 +
              1*Var.5
              
linearSlope =~ 0*Var.1 +
              1*Var.2 +
              2*Var.3 +
              3*Var.4 +
              4*Var.5
              
'
```

## Piecewise

```{r}
piecewise.model.generic <- '

myIntercept =~ 1*Var.1 +
              1*Var.2 +
              1*Var.3 +
              1*Var.4 +
              1*Var.5
              
slope1 =~ 0*Var.1 +
              1*Var.2 +
              2*Var.3 +
              2*Var.4 +
              2*Var.5
              
slope2 =~ 0*Var.1 +
              0*Var.2 +
              0*Var.3 +
              1*Var.4 +
              2*Var.5
              
'
```

## Moderated Piecewise

```{r}

# function fills "ageGeneric" with a named variable from the dataset
# fills to ageMeanCentered by default

piecewise.model.moderated.generic <- '

myIntercept =~ 1*Var.1 +
              1*Var.2 +
              1*Var.3 +
              1*Var.4 +
              1*Var.5
              
slope1 =~ 0*Var.1 +
              1*Var.2 +
              2*Var.3 +
              2*Var.4 +
              2*Var.5
              
slope2 =~ 0*Var.1 +
              0*Var.2 +
              0*Var.3 +
              1*Var.4 +
              2*Var.5

slope1 ~~ slope1Sigma2*slope1
slope2 ~~ slope2Sigma2*slope2
              
myIntercept ~ mIntercept*ageGeneric
slope1 ~ mSlope1*ageGeneric
slope2 ~ mSlope2*ageGeneric

ageES_slope1 := mSlope1/sqrt(slope1Sigma2)
ageES_slope2 := mSlope2/sqrt(slope2Sigma2)

              
'

```

```{r}

fitLinearModel <- function(instrumentName){
  # input instrumentName as argument
  
  
  model <- str_replace_all(linear.model.generic, 
                           c("Var" = instrumentName))
  
  fit <- lavaan::growth(model = model,
                        data = data.wide,
                        missing = 'FIML')
  
  return(fit)
  
}

```

```{r}

fitPiecewiseModel <- function(instrumentName){
  # input instrumentName as argument
  
  model <- str_replace_all(
                    piecewise.model.generic, 
                    c("Var" = instrumentName))
  
  fit <- lavaan::growth(model = model,
                        data = data.wide,
                        missing = 'FIML')
  
  return(fit)
  
}

```

```{r}

fitPiecewiseModeratedModel <- function(instrumentName, 
                                       ageVar = "ageMeanCentered" ){
  # input instrumentName as argument
  
  
  model <- str_replace_all(piecewise.model.moderated.generic, c("Var" = instrumentName,
                                                                "ageGeneric" = ageVar))
  
  fit <- lavaan::growth(model = model,
                        data = data.wide,
                        missing = 'FIML')
  
  return(fit)
  
}

```

# Fit models to data

## Pain severity

### Linear

```{r}

BPIseverity.linear.fit <- fitLinearModel('BPIseverity')

summary(BPIseverity.linear.fit, fit.measures = TRUE)

```

### Piecewise

```{r}

BPIseverity.piecewise.fit <- fitPiecewiseModel('BPIseverity')

summary(BPIseverity.piecewise.fit, fit.measures = TRUE)

```

### Model comparison

```{r}
anova(BPIseverity.linear.fit, BPIseverity.piecewise.fit)

```

### Moderated piecewise

```{r}

BPIseverity.piecewise.moderated.fit <- fitPiecewiseModeratedModel('BPIseverity')

summary(BPIseverity.piecewise.moderated.fit, fit.measures = TRUE)

write_rds(x = BPIseverity.piecewise.moderated.fit, 
          file = here('Model fits', 'BPIseverity.piecewise.moderated.fit'))

```

### Other moderated models (Pick-A-Point)

```{r}

BPIseverity.centered35.fit <-
  fitPiecewiseModeratedModel('BPIseverity', ageVar = 'ageCenteredAt35')

BPIseverity.centered45.fit <-
  fitPiecewiseModeratedModel('BPIseverity', ageVar = 'ageCenteredAt45')

BPIseverity.centered55.fit <-
  fitPiecewiseModeratedModel('BPIseverity', ageVar = 'ageCenteredAt55')

BPIseverity.centered65.fit <-
  fitPiecewiseModeratedModel('BPIseverity', ageVar = 'ageCenteredAt65')


write_rds(x = BPIseverity.centered35.fit, 
          file = here('Model fits', 'Pick a point',
                      'BPIseverity.centered35.fit'))


write_rds(x = BPIseverity.centered45.fit, 
          file = here('Model fits', 'Pick a point',
                      'BPIseverity.centered45.fit'))

write_rds(x = BPIseverity.centered55.fit, 
          file = here('Model fits', 'Pick a point',
                      'BPIseverity.centered55.fit'))

write_rds(x = BPIseverity.centered65.fit, 
          file = here('Model fits', 'Pick a point',
                      'BPIseverity.centered65.fit'))

```

## Pain interference

### Linear

```{r}

BPIinterference.linear.fit <- fitLinearModel('BPIinterference')

summary(BPIinterference.linear.fit, fit.measures = TRUE)

```

### Piecewise

```{r}

BPIinterference.piecewise.fit <- fitPiecewiseModel('BPIinterference')

summary(BPIinterference.piecewise.fit, fit.measures = TRUE)

```

### Model comparison

```{r}
anova(BPIinterference.linear.fit, BPIinterference.piecewise.fit)

```

### Moderated piecewise

```{r}

BPIinterference.piecewise.moderated.fit <- fitPiecewiseModeratedModel('BPIinterference')

summary(BPIinterference.piecewise.moderated.fit, fit.measures = TRUE)

```

### Other moderated models (Pick-A-Point)

```{r}

BPIinterference.centered35.fit <-
  fitPiecewiseModeratedModel('BPIinterference', ageVar = 'ageCenteredAt35')

BPIinterference.centered45.fit <-
  fitPiecewiseModeratedModel('BPIinterference', ageVar = 'ageCenteredAt45')

BPIinterference.centered55.fit <-
  fitPiecewiseModeratedModel('BPIinterference', ageVar = 'ageCenteredAt55')

BPIinterference.centered65.fit <-
  fitPiecewiseModeratedModel('BPIinterference', ageVar = 'ageCenteredAt65')


write_rds(x = BPIinterference.centered35.fit, 
          file = here('Model fits', 'Pick a point',
                      'BPIinterference.centered35.fit'))


write_rds(x = BPIinterference.centered45.fit, 
          file = here('Model fits', 'Pick a point',
                      'BPIinterference.centered45.fit'))

write_rds(x = BPIinterference.centered55.fit, 
          file = here('Model fits', 'Pick a point',
                      'BPIinterference.centered55.fit'))

write_rds(x = BPIinterference.centered65.fit, 
          file = here('Model fits', 'Pick a point',
                      'BPIinterference.centered65.fit'))

```

## CPAQ total

Piecewise is best fit though negative resid var at T5, no age moderation effects

### Linear

```{r}

CPAQ.linear.fit <- fitLinearModel('CPAQ')

summary(CPAQ.linear.fit, fit.measures = TRUE)

```

### Piecewise

Timepoint 5 observed variable variance \< 0, p = ns Thus re-run with that variance set to zero below

```{r}

CPAQ.piecewise.fit <- fitPiecewiseModel('CPAQ')

summary(CPAQ.piecewise.fit, fit.measures = TRUE)

```

### Piecewise, timepoint 5 variance set to zero

See above

```{r}

CPAQ.piecewise.T5varIsZero.model <- '

myIntercept =~ 1*CPAQ.1 +
              1*CPAQ.2 +
              1*CPAQ.3 +
              1*CPAQ.4 +
              1*CPAQ.5
              
slope1 =~ 0*CPAQ.1 +
              1*CPAQ.2 +
              2*CPAQ.3 +
              2*CPAQ.4 +
              2*CPAQ.5
              
slope2 =~ 0*CPAQ.1 +
              0*CPAQ.2 +
              0*CPAQ.3 +
              1*CPAQ.4 +
              2*CPAQ.5
              
              
CPAQ.5 ~~ 0*CPAQ.5

slope1 ~~ slope1Sigma2*slope1
slope2 ~~ slope2Sigma2*slope2
              
# myIntercept ~ mIntercept*ageMeanCentered
# slope1 ~ mSlope1*ageMeanCentered
# slope2 ~ mSlope2*ageMeanCentered
'

CPAQ.piecewise.T5varIsZero.fit <- lavaan::growth(model = CPAQ.piecewise.T5varIsZero.model, 
                                                 data = data.wide,
                        missing = 'FIML')


summary(CPAQ.piecewise.T5varIsZero.fit, fit.measures = TRUE)

```

### Model comparison

```{r}
anova(CPAQ.linear.fit, CPAQ.piecewise.fit)

anova(CPAQ.linear.fit, CPAQ.piecewise.T5varIsZero.fit)

```

### Moderated piecewise

```{r}

CPAQ.piecewise.moderated.model <-'

myIntercept =~ 1*CPAQ.1 +
              1*CPAQ.2 +
              1*CPAQ.3 +
              1*CPAQ.4 +
              1*CPAQ.5
              
slope1 =~ 0*CPAQ.1 +
              1*CPAQ.2 +
              2*CPAQ.3 +
              2*CPAQ.4 +
              2*CPAQ.5
              
slope2 =~ 0*CPAQ.1 +
              0*CPAQ.2 +
              0*CPAQ.3 +
              1*CPAQ.4 +
              2*CPAQ.5
              
              
CPAQ.5 ~~ 0*CPAQ.5

slope1 ~~ slope1Sigma2*slope1
slope2 ~~ slope2Sigma2*slope2
              
myIntercept ~ mIntercept*ageMeanCentered
slope1 ~ mSlope1*ageMeanCentered
slope2 ~ mSlope2*ageMeanCentered'

CPAQ.piecewise.moderated.fit <- lavaan::growth(model = CPAQ.piecewise.moderated.model, 
                                                 data = data.wide,
                        missing = 'FIML')

summary(CPAQ.piecewise.moderated.fit, fit.measures = TRUE)

```

## CPAQ Activities

Piecewise is best fit, no age moderation effects

### Linear

```{r}

CPAQ.linear.fit <- fitLinearModel('CPAQ_Activities')

summary(CPAQ.linear.fit, fit.measures = TRUE)

```

### Piecewise

```{r}

CPAQ.piecewise.fit <- fitPiecewiseModel('CPAQ_Activities')

summary(CPAQ.piecewise.fit, fit.measures = TRUE)

```

### Model comparison

```{r}
anova(CPAQ.linear.fit, CPAQ.piecewise.fit)

```

### Moderated piecewise

```{r}

CPAQ.piecewise.moderated.fit <- fitPiecewiseModeratedModel('CPAQ_Activities')

summary(CPAQ.piecewise.moderated.fit, fit.measures = TRUE)

```

## CPAQ Willingness

Piecewise is best fit though negative resid var at T5, no age moderation effects

### Linear

```{r}

CPAQ.linear.fit <- fitLinearModel('CPAQ_Willingness')

summary(CPAQ.linear.fit, fit.measures = TRUE)

```

### Piecewise

```{r}

CPAQ.piecewise.fit <- fitPiecewiseModel('CPAQ_Willingness')

summary(CPAQ.piecewise.fit, fit.measures = TRUE)

```

### Model comparison

```{r}
anova(CPAQ.linear.fit, CPAQ.piecewise.fit)

```

### Moderated piecewise

```{r}

CPAQ.piecewise.moderated.fit <- fitPiecewiseModeratedModel('CPAQ_Willingness')

summary(CPAQ.piecewise.moderated.fit, fit.measures = TRUE)

```

## PHQ9

Piecewise is best fit though negative resid var at T5, no age moderation effects

### Linear

```{r}

PHQ9.linear.fit <- fitLinearModel('PHQ9')

summary(PHQ9.linear.fit, fit.measures = TRUE)

```

### Piecewise

```{r}

PHQ9.piecewise.fit <- fitPiecewiseModel('PHQ9')

summary(PHQ9.piecewise.fit, fit.measures = TRUE)

```

### Model comparison

```{r}
anova(PHQ9.linear.fit, PHQ9.piecewise.fit)

```

### Moderated piecewise

```{r}

PHQ9.piecewise.moderated.fit <- fitPiecewiseModeratedModel('PHQ9')

summary(PHQ9.piecewise.moderated.fit, fit.measures = TRUE)

```

# PSQI

### Linear

```{r}

PSQI.linear.fit <- fitLinearModel('PSQI')

summary(PSQI.linear.fit, fit.measures = TRUE)

```

### Piecewise

```{r}

PSQI.piecewise.fit <- fitPiecewiseModel('PSQI')

summary(PSQI.piecewise.fit, fit.measures = TRUE)

```

### Model comparison

```{r}
anova(PSQI.linear.fit, PSQI.piecewise.fit)

```

### Moderated linear

```{r}

PSQI.linear.moderated.model <-'

myIntercept =~ 1*PSQI.1 +
              1*PSQI.2 +
              1*PSQI.3 +
              1*PSQI.4 +
              1*PSQI.5
              
linearSlope =~ 0*PSQI.1 +
              1*PSQI.2 +
              2*PSQI.3 +
              3*PSQI.4 +
              4*PSQI.5
              
myIntercept ~ mIntercept*ageMeanCentered
linearSlope ~ mSlope1*ageMeanCentered

              
'

PSQI.linear.moderated.fit <- lavaan::growth(model = PSQI.linear.moderated.model,
                        data = data.wide,
                        missing = 'FIML')

summary(PSQI.linear.moderated.fit, fit.measures = TRUE)

```

# Save models

## Unmoderated

```{r}

write_rds(x = BPIinterference.piecewise.fit,file = 
            here('Model fits',
                 'Unmoderated',
                 'BPIinterference.piecewise.fit'))

write_rds(x = BPIseverity.piecewise.fit,file = 
            here('Model fits',
                 'Unmoderated',
                 'BPIseverity.piecewise.fit'))

write_rds(x = CPAQ.piecewise.T5varIsZero.fit,file = 
            here('Model fits',
                 'Unmoderated',
                 'CPAQ.piecewise.fit'))

write_rds(x = PHQ9.piecewise.fit,file = 
            here('Model fits',
                 'Unmoderated',
                 'PHQ9.piecewise.fit'))

write_rds(x = PSQI.linear.fit,file = 
            here('Model fits',
                 'Unmoderated',
                 'PSQI.linear.fit'))


```

## Moderated

```{r}

write_rds(x = BPIinterference.piecewise.moderated.fit,file = 
            here('Model fits',
                 'Moderated',
                 'BPIinterference.piecewise.moderated.fit'))

write_rds(x = BPIseverity.piecewise.moderated.fit,file = 
            here('Model fits',
                 'Moderated',
                 'BPIseverity.piecewise.moderated.fit'))

write_rds(x = CPAQ.piecewise.moderated.fit,file = 
            here('Model fits',
                 'Moderated',
                 'CPAQ.piecewise.moderated.fit'))

write_rds(x = PHQ9.piecewise.moderated.fit,file = 
            here('Model fits',
                 'Moderated',
                 'PHQ9.piecewise.moderated.fit'))

write_rds(x = PSQI.linear.moderated.fit,file = 
            here('Model fits',
                 'Moderated',
                 'PSQI.linear.moderated.fit'))


```

## Pick a point

```{r}

write_rds(x = BPIseverity.centered35.fit,file = 
            here('Model fits',
                 'Moderated',
                 'Pick-A-Point',
                 'BPIseverity.centered35.fit'))

write_rds(x = BPIseverity.centered45.fit,file = 
            here('Model fits',
                 'Moderated',
                 'Pick-A-Point',
                 'BPIseverity.centered45.fit'))

write_rds(x = BPIseverity.centered55.fit,file = 
            here('Model fits',
                 'Moderated',
                 'Pick-A-Point',
                 'BPIseverity.centered55.fit'))

write_rds(x = BPIseverity.centered65.fit,file = 
            here('Model fits',
                 'Moderated',
                 'Pick-A-Point',
                 'BPIseverity.centered65.fit'))


write_rds(x = BPIinterference.centered35.fit,file = 
            here('Model fits',
                 'Moderated',
                 'Pick-A-Point',
                 'BPIinterference.centered35.fit'))

write_rds(x = BPIinterference.centered45.fit,file = 
            here('Model fits',
                 'Moderated',
                 'Pick-A-Point',
                 'BPIinterference.centered45.fit'))

write_rds(x = BPIinterference.centered55.fit,file = 
            here('Model fits',
                 'Moderated',
                 'Pick-A-Point',
                 'BPIinterference.centered55.fit'))

write_rds(x = BPIinterference.centered65.fit,file = 
            here('Model fits',
                 'Moderated',
                 'Pick-A-Point',
                 'BPIinterference.centered65.fit'))

```
