---
title: "Explore data"
author: "Joel Fishbein"
format: html
editor: visual
---

# Read in data and libraries

```{r}
library(here)
library(tidyverse)

data.allvars <- read_csv(here('TeleACT-Aging_Joel_9-16-22.csv'))

```

# Examine variables of interest

-   ID - participant ID

-   Telehealth: participant's randomized condition assignment of Telehealth ( = ?) vs in -person ACT (= ?)

-   age - age at baseline in years?

-   Age_Bi: age variable dichotomized by Matt at age ??

-   BPIinterference.1-BPIinterference.5: pain interference

-   BPIseverity.1 - BPIseverity.5: pain severity

-   CPAQ.1-CPAQ.5: chronic pain acceptance total score?

-   CPAQ_Activities.1-CPAC_Activities.5: chronic pain acceptance activities score?

-   CPAQ_Willingness.1-CPAC_Willingness.5: chronic pain acceptance activities score?

```{r}

colnames(data.allvars)
```

# Select variables to examine

```{r}



data.wide <- data.allvars %>%
  select(
    starts_with(
      c("ID", "Telehealth", "age", "Age_Bi",
        "BPIinterference", "BPIseverity",
        "CPAQ.", "CPAQ_Activities", "CPAQ_Willingness")))

colnames(data.wide)


```

# Pivot long

```{r}

data.long <- data.wide %>%
  pivot_longer(cols = BPIinterference.1:CPAQ_Willingness.5,
               names_to = c("instrument", "timepoint"), names_pattern = "(.*).(.)")


```
